---
title: 'Class Exercise 1: 4773'
author: "Collin Dabbieri"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Make an R function
This function will be entirely your own work - you can however use other packages to create the plot if you wish.

## Make an R function that will do the following
Look at the formula MS page 446.
For a one way table:

>
  1. Create point estimates for $p_i-p_j$
  2. Create ci's for $p_i-p_j$
  3. Create a colored barplot.
  
```{r mypipj}
mypipj = function(tab,ni,nj,alpha=0.05) #tab = table
{
  total_points=sum(tab)
  ci_matrix_lower=array(dim=c(length(tab),length(tab)))
  ci_matrix_upper=array(dim=c(length(tab),length(tab)))
  pt_est_matrix=array(dim=c(length(tab),length(tab)))
  z=qnorm(1-alpha/2,0,1)
  for (i in 1:length(tab)){
    for (j in 1:length(tab)){
      if (i==j){
        
      }
      else{
        pihat=tab[i]/total_points
        pjhat=tab[j]/total_points
        pt_est=pihat-pjhat
        pt_est_matrix[i,j]=pt_est
      
        radicand=(pihat*(1-pihat)+pjhat*(1-pjhat)+2*pihat*pjhat)/(total_points)
        lower=pt_est-z*sqrt(radicand)
        upper=pt_est+z*sqrt(radicand)
        ci=c(lower,upper)
        ci_matrix_lower[i,j]=lower
        ci_matrix_upper[i,j]=upper
        
      }

      
    }
    
  }
  
  dimnames(ci_matrix_lower) = list( c("A", "B","C","D","E"),c("A","B","C","D","E"))
  dimnames(ci_matrix_upper) = list( c("A", "B","C","D","E"),c("A","B","C","D","E")) 
  dimnames(pt_est_matrix) = list( c("A", "B","C","D","E"),c("A","B","C","D","E"))

  
  
  # Now create colored barplot with hypothesis values and sample values
  num_categories=length(tab)
  hypoth_val=1./num_categories
 
  
  true_vals=c()
  for (i in 1:num_categories){
    ptemp=tab[i]/total_points
    true_vals=append(true_vals,ptemp)
  }

  
  barplot(rbind(true_vals,hypoth_val),beside=TRUE,legend = c("sample", "hypothesis"),xlab="Categories",ylab="Proportion",col=c('lightblue','magenta'))
  
  return(list(ci_lower_bound=ci_matrix_lower,ci_upper_bound=ci_matrix_upper,pt_est=pt_est_matrix))
  
}
```


## Use data in your function

Use the IMPELLER.xls (see MS pg 444) file for the table.

```{r}
library(readxl)
IMPELLER <- read_excel("/home/collindabbieri/Documents/AppliedRegressionAnalysis/Dataxls/Excel/IMPELLER.XLS")
#head(IMPELLER,font="courier new")
f=IMPELLER$DEFECTIVE
nf=IMPELLER$LINE
m=rep(nf,f)
imptab=table(m)
imptab

```


```{r impeller}
list=mypipj(imptab,ni=31,nj=11,alpha=0.05)
list
```

