---
title: "Laboratory 5"
author: "Collin Dabbieri"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# My functions {.tabset .tabset-pills .tabset-fade}

## myy()
```{r}
myy=function(Y,X){
  n=nrow(X)
  k=ncol(X)-1
  model=lm(Y~X[,-1])
  r=model$residuals
  coeff=model$coefficients

  SSE=sum(r^2)
  s=sqrt(SSE/(n-(k+1)))
  t=qt(1-alpha/2,n-(k+1))
  pm=c(-1,1)
  
  ci_matrix=matrix(,nrow=n,ncol=2)
  yhat_vector=c()
  for(i in 1:n){
    x=X[i,]

    yhat=0
    for(j in 1:k+1){
      yhat=yhat+(coeff[j]*x[j])
    }

    ci=yhat+pm*t*s*sqrt(1+t(x)%*%solve(t(X)%*%X)%*%x)
    ci_matrix[i,1]=ci[1]
    ci_matrix[i,2]=ci[2]
    yhat_vector=append(yhat_vector,yhat)

  }
  
  
  
  ##### Pairs Plot
  
  pairs_data=X
  pairs_data[,1]=Y

  pairs(pairs_data)
  
  
  
  ###### ggplot of Y vs xi
  # I don't understand what you're asking for here
  
  
  ###### plot of cis and point estimates (x=1:n)
  index=seq(n)
  plot(index,yhat_vector,ylim=c(min(ci_matrix[,1]),max(ci_matrix[,2])))
  for(i in 1:n){
    segments(x0=i,y0=ci_matrix[i,1],y1=ci_matrix[i,2])
  }
  
  
  names(yhat_vector)=seq(n)
  return(list(yhat=yhat_vector,ci=ci_matrix))
}
```


## myl()

```{r}
myl=function(Y,X,alpha,a,l){
  
  n=nrow(X)
  k=ncol(X)-1
  t=qt(1-alpha/2,n-(k+1))
  pm=c(-1,1)
  
  model=lm(Y~X[,-1])
  coeff=model$coefficients
  
  betahat=matrix(data=coeff,nrow=k+1,ncol=1)
  
  coeff=model$coefficients
  r=model$residuals
  SSE=sum(r^2)
  s=sqrt(SSE/(n-(k+1)))
  
  ci=l+pm*t*s*sqrt(t(a)%*%solve(t(X)%*%X)%*%a)

  pairs(X)
  return(list(ci=ci))
  
}
```


# Task 1

```{r}
getwd()
```


# Task 2 {.tabset .tabset-pills .tabset-fade}

```{r}
library(readxl)

clerical=read_excel("../../Dataxls/Excel/CLERICAL.XLS")
head(clerical)
```

Replace the first 4 entries in the column "X5-Checks" with c(600,500,1000,900) 

```{r}
print(clerical$`X5-Checks`[1:4])
clerical$`X5-Checks`[1:4]=c(600,500,1000,900)
print(clerical$`X5-Checks`[1:4])
```

## A. 
Use scattergrams to plot the sample data. Interpret the results.

```{r}
clerical_plot=subset(clerical,select=-c(Obs,Day))
pairs(clerical_plot)
```


## B
Use the method of Least Squares to estimate the model parameters. Interpret the $\beta$-estimates.

```{r}
n=length(clerical$`Y-Hours`)
k=7

model=lm(`Y-Hours`~.,data=clerical_plot)
summary(model)
```


A few of the $\beta$ parameters seem important at the start (X4 and X5)

## C
Find the estimate of $\sigma$, the standard deviation of the random error term, and interpret its value.


```{r}
coeff=model$coefficients
y_fit=coeff[1]+coeff[2]*clerical$`X1-Mail`+coeff[3]*clerical$`X2-Gifts`+coeff[4]*clerical$`X3-Charge`+coeff[5]*clerical$`X4-Returns`+coeff[6]*clerical$`X5-Checks`+coeff[7]*clerical$`X6-Misc`+coeff[8]*clerical$`X7-Tickets`
y=clerical$`Y-Hours`
r=y-y_fit
SSE=sum(r^2)
s=sqrt(SSE/(n-(k+1)))
s

```


## D
Assess the adequcy of the model by conducting a test of hypothesis at $\alpha=0.05$

```{r}
summary(model)
```
The NULL for the F test is $H_0:\beta_1=\beta_2=...=\beta_k=0$. We see evidence against this NULL at the 0.05 level. The model is adequate. At least one of the model parameters is nonzero at the 95% confidence level.

## E
Find a 95% confidence interval for $\beta_2$. Interpret the result.

```{r}
X=matrix(,nrow=n,ncol=k+1)
X[,1]=1
X[,2]=clerical$`X1-Mail`
X[,3]=clerical$`X2-Gifts`
X[,4]=clerical$`X3-Charge`
X[,5]=clerical$`X4-Returns`
X[,6]=clerical$`X5-Checks`
X[,7]=clerical$`X6-Misc`
X[,8]=clerical$`X7-Tickets`

Y=matrix(,nrow=n,ncol=1)
Y[,1]=clerical$`Y-Hours`

betahat=matrix(coeff,nrow=k+1,ncol=1)
a=matrix(c(0,0,1,0,0,0,0,0,0),nrow=k+1,ncol=1)
pm=c(-1,1)
alpha=0.05

ci=t(a)%*%betahat+pm*qt(1-alpha/2,n-(k+1))*s*sqrt(t(a)%*%solve(t(X)%*%X)%*%a)
coeff[2]
ci
```
We are 95% confident that $\beta_2$ falls within that confidence interval.

## F
Find the adjusted coefficient of determination, $R_a^2$, and interpret the result.

```{r}
summary(model)
```
The Adjusted $R^2$ is 0.49. This value can be used to compare this model with other models of varying complexity because it includes a penalty for complexity of the model.

## G
Find a 95% prediction interval for the number of hours worked on a day when $x_1=5000$ pieces of mail processed, $x_2=75$ gift certificates sold, $x_3=900$ store charge accounts transactions made, $x_4=200$ change order transactions processed, and $x_5=650$ checks cashed. Interpret the result.

```{r}
a=matrix(c(1,5000,75,900,200,650,0,0,0),nrow=k+1,ncol=1)
ci=t(a)%*%betahat+pm*qt(1-alpha/2,n-(k+1))*s*sqrt(t(a)%*%solve(t(X)%*%X)%*%a)
ci

```
We can be 95% confident that the number of hours worked on a day with the given specifications will be between 85.3 and 122.6

# Task 3

Using Theorem 11.3 page 574, make a function myl() that produces

-(1-$\alpha$)100% confidence intervals for E(l) given Y, X, $\alpha$, a, l
-a plot using pairs() to show all pairwise plots of the data used in the analysis

```{r mylfunction}
myl=function(Y,X,alpha,a,l){
  
  n=nrow(X)
  k=ncol(X)-1
  t=qt(1-alpha/2,n-(k+1))
  pm=c(-1,1)
  
  model=lm(Y~X[,-1])
  coeff=model$coefficients
  
  betahat=matrix(data=coeff,nrow=k+1,ncol=1)
  
  coeff=model$coefficients
  r=model$residuals
  SSE=sum(r^2)
  s=sqrt(SSE/(n-(k+1)))
  
  ci=l+pm*t*s*sqrt(t(a)%*%solve(t(X)%*%X)%*%a)

  pairs(X)
  return(list(ci=ci))
  
}
```


Using the modified CLERICAL data with $\alpha=0.05$, and a such that $E(l)=\beta_1-\beta_2$ call your function and place output here

```{r ci_intext}
X=matrix(,nrow=n,ncol=k+1)
X[,1]=1
X[,2]=clerical$`X1-Mail`
X[,3]=clerical$`X2-Gifts`
X[,4]=clerical$`X3-Charge`
X[,5]=clerical$`X4-Returns`
X[,6]=clerical$`X5-Checks`
X[,7]=clerical$`X6-Misc`
X[,8]=clerical$`X7-Tickets`

Y=matrix(,nrow=n,ncol=1)
Y[,1]=clerical$`Y-Hours`

a=matrix(c(0,1,-1,0,0,0,0,0),nrow=k+1,ncol=1)

model=lm(Y~X[,-1])
coeff=model$coefficients
betahat=matrix(coeff,nrow=k+1,ncol=1)
l=t(a)%*%betahat

alpha=0.05
ci=myl(Y,X,alpha,a,l)
ci

```

Interpret the confidence interval that your function produces.

We can say with 95% confidence that $\beta_1-\beta_2$ will fall in this range.

# Task 5

Find a 95% ci for E(Y) when the x variables take the second row of the design matrix


What is the exact expression for a?

```{r}
a=matrix(data=c(0,1,0,0,0,0,0,0),nrow=8,ncol=1)
a
```


Use your function and page 575, place output here

```{r}
ci=myl(Y,X,alpha,a,l)
ci
```

Interpret the confidence interval - be careful!


# Task 6
Make a confidence interval for a new y value
```{r}
#x is a vector with values of x for the new y value
myy=function(x,Y,X){
  n=nrow(X)
  k=ncol(X)-1
  model=lm(Y~X[,-1])
  r=model$residuals
  coeff=model$coefficients
  yhat=coeff[1]
  for(i in 1:k){
    yhat=yhat+(coeff[i+1]*x[i])
  }

  SSE=sum(r^2)
  s=sqrt(SSE/(n-(k+1)))
  t=qt(1-alpha/2,n-(k+1))
  pm=c(-1,1)
  
  ci=yhat+pm*t*s*sqrt(1+t(x)%*%solve(t(X)%*%X)%*%x)
  
  return(list(yhat=yhat,ci=ci))
}
```

use the second row of the design matrix to predict a new y value and get a ci for the new y value

```{r}
x=matrix(X[2,],nrow=k+1,ncol=1)
myy(x,Y,X)
```


Alter the function myy() so it produces
-pairs plot
-ggplot of Y vs xi for each i
-a plot of the cis and point estimates (x=1:n)
-a list containing the ci and point estimate for a new y value

```{r}
myy=function(Y,X){
  n=nrow(X)
  k=ncol(X)-1
  model=lm(Y~X[,-1])
  r=model$residuals
  coeff=model$coefficients

  SSE=sum(r^2)
  s=sqrt(SSE/(n-(k+1)))
  t=qt(1-alpha/2,n-(k+1))
  pm=c(-1,1)
  
  ci_matrix=matrix(,nrow=n,ncol=2)
  yhat_vector=c()
  for(i in 1:n){
    x=X[i,]

    yhat=0
    for(j in 1:k+1){
      yhat=yhat+(coeff[j]*x[j])
    }

    ci=yhat+pm*t*s*sqrt(1+t(x)%*%solve(t(X)%*%X)%*%x)
    ci_matrix[i,1]=ci[1]
    ci_matrix[i,2]=ci[2]
    yhat_vector=append(yhat_vector,yhat)

  }
  
  
  
  ##### Pairs Plot
  
  pairs_data=X
  pairs_data[,1]=Y

  pairs(pairs_data)
  
  
  
  ###### ggplot of Y vs xi
  # I don't understand what you're asking for here
  
  
  ###### plot of cis and point estimates (x=1:n)
  index=seq(n)
  plot(index,yhat_vector,ylim=c(min(ci_matrix[,1]),max(ci_matrix[,2])))
  for(i in 1:n){
    segments(x0=i,y0=ci_matrix[i,1],y1=ci_matrix[i,2])
  }
  
  
  names(yhat_vector)=seq(n)
  return(list(yhat=yhat_vector,ci=ci_matrix))
}
```


```{r}
options(warn=-1)
list=myy(Y,X)
list
```

# Task 7

Add the two functions to the ILAS2019 package

```{r}
library(ILAS2019)

ILAS2019::myl(Y,X,alpha,a,l)
ILAS2019::myy(Y,X)
```






