---
title: "Laboratory 5"
author: "Collin Dabbieri"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# My functions {.tabset .tabset-pills .tabset-fade}

## Function 1

## Function 2


# Task 1

```{r}
getwd()
```


# Task 2 {.tabset .tabset-pills .tabset-fade}

```{r}
library(readxl)

clerical=read_excel("../../Dataxls/Excel/CLERICAL.XLS")
head(clerical)
```

Replace the first 4 entries in the column "X5-Checks" with c(600,500,1000,900) 

```{r}
print(clerical$`X5-Checks`[1:4])
clerical$`X5-Checks`[1:4]=c(600,500,1000,900)
print(clerical$`X5-Checks`[1:4])
```

## A

## B

## C

## D

## E

## F

## G

# Task 3

Using Theorem 11.3 page 574, make a function myl() that produces

-(1-$\alpha$)100% confidence intervals for E(l) given Y, X, $\alpha$, a, l
-a plot using pairs() to show all pairwise plots of the data used in the analysis

```{r mylfunction}
myl=function(Y,X,alpha,a,l){
  
  #a'betahat +- t s sqrt(a'(X'X)^-1 a)
  betahat=
  
}
```


Using the modified CLERICAL data with $\alpha=0.05$, and a such that $E(l)=\beta_1-\beta_2$ call your function and place output here

```{r}
Y=
X=
a=
alpha=0.05
betahat=
l=t(a)%*%betahat
ci=myl(Y,X,alpha,a,l)
```


Interpret the confidence interval that your function produces

The confidence interval is `r ci`

# Task 5

Find a 95% ci for E(Y) when the x variables take the second row of the design matrix


What is the exact expression for a?

```{r}
a=matrix(data=c(0,1,0,0,0,0,0,0,0),nrow=9,ncol=1)
a
```


Use your function and page 575, place output here

```{r}
ci=myl(Y,X,alpha,a,l)
```

Interpret the confidence interval - be careful!


# Task 6





