---
title: "Laboratory 11"
author: "Collin Dabbieri"
date: "11/11/2019"
output: 
  html_document:
    toc: yes
    toc_float: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1

```{r}
getwd()
```

# Task 2

Using the package s20x load the data arousal.df

```{r}
library(s20x)

data(arousal.df)
head(arousal.df)
```


Write a paragraph on what the data and experiment concerns.

Emotional arousal levels were quantified for males and females by measuring changes in their pupil diameters. Participants were presented an image of either a nude male, a nude female, a landscape or an infant and their arousal levels were recorded. 20 males and 20 females participated in the study.


Make an interaction plot with gender on the horizontal axis

```{r}
interactionPlots(arousal~gender+picture,data=arousal.df)
```



Do the same but with picture on the horizontal axis

```{r}
interactionPlots(arousal~picture+gender,data=arousal.df)
```


What do you conclude from the pictures?

The pictures tell me that gender and picture clearly interact, so a gender:picture term will be necessary in the model.


Give the anova table for the interactive model

```{r}
model=lm(arousal~gender+picture+gender:picture,data=arousal.df)
summary(model)
anova(model)
```

Interpret the output

Our model is

$$E(y)=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_3+\beta_4x_4+\beta_5x_1x_2+\beta_6x_1x_3+\beta_7x_1x_4$$

Where $x_1=1$ when gender is male. $x_2=1$ when picture is landscape. $x_3=1$ when picture is nude female and $x_4=1$ when picture is nude male. So female gender and infant picture are our baselines.

The anova table tells us that the gender terms, the picture terms, and the interaction terms are significant in our model at the 95% confidence level.


Give the point and 95% interval estimate for male:landscape-female:landscape


```{r}
gender.picture<-crossFactors(arousal.df$gender,arousal.df$picture)
gender.picture<-factor(gender.picture)
arousal.1way.fit<-lm(arousal.df$arousal~gender.picture)
summary1way(arousal.1way.fit)
multipleComp(arousal.1way.fit)
```

We can see from the output that the female:landscape-male:landscape has estimate 0.885 with 95% confidence interval (-0.043-1.813)

So male:landscape-female:landscape has estimate -0.885 with confidence interval (-1.813-0.043)

Interpret the  interval

We can see from the interaction plot that males and females seem to have similar arousal levels when presented with a picture of a landscape. This is confirmed by the interval, we can say it is plausible at the 95% confidence level that males and females have the same arousal reaction to a landscape.