---
title: "Exam 2 Answer Booklet"
author: "Dr Wayne Stewart"
date: "11/8/2019"
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Name: Collin Dabbieri

Grade 2,3,4 please

Please place answers to question 2-4 here: 
\newpage

# Question 2

```{r}
library(readxl)
prodqual=read_excel("./PRODQUAL.XLS")
head(prodqual)
```


## a)

$$E(y)=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_1x_2+\beta_4x_1^2+\beta_5x_2^2$$

## b)

```{r}
qual=prodqual$QUALITY
temp=prodqual$TEMP
pressure=prodqual$PRESSURE
temp2=temp^2
press2=pressure^2
model=lm(qual~temp+pressure+temp:pressure+temp2+press2)
summary(model)
```


## c)

The p-value from the F-test (2.2e-16) tells us that the model is adequate at the 95% confidence level.

## d)

99.3% of the variability in the response is explained by the model. This comes from the $R^2$ value

$$R^2=1-\frac{SSE}{TSS}$$

## e)

$$y=-5.128e3+3.11e1x_1+1.397e2x_2-1.334e-1x_3-1.144e0x_4-1.455e-1x_5$$

Maximum point has a slope of 0

$$y'=0=$$

```{r}
A=matrix(c(-5.128e3,3.11e1,1.397e2,-1.334e-1,-1.144e0,-1.455e-1))
```



\newpage

# Question 3

```{r}
library(s20x)

```


## a)

```{r}
data(seeds.df)
head(seeds.df)
seeds.df=within(seeds.df,{Water.F=factor(Water)})
model1=lm(Count~Light+Water.F+Light:Water.F,data=seeds.df)
eovcheck(model1)
```


There does not appear to be fanning, I believe we can say that it is plausible that the variance is constant.





## b)

```{r}
interactionPlots(Count~Light+Water.F,data=seeds.df)
interactionPlots(Count~Water.F+Light,data=seeds.df)
```

There is some pretty strong evidence for interaction here. The lines are certainly not parallel.

## c)

```{r}
summary2way(model1)
```

the p-value for the interaction term is ~0. We have strong evidence for interaction.


## d)

The main effects appear to be water and the interaction term. The light term is not significant at the 95% confidence level.


## e)

The most parsiminous model possible will only use either Water or Light

```{r}
model2=lm(Count~Water.F+Light:Water.F,data=seeds.df)
anova(model2)
```
Let's remove the interaction term

```{r}
model3=lm(Count~Water.F,data=seeds.df)
summary(model3)
```

So we're left with only Water.


Water has 6 levels, 5 dummy variables


$$E(y)=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_3+\beta_4x_4+\beta_5x_5$$



\newpage

# Question 4

## a)

$$E(y)=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_3+\beta_4x_4$$

the $x_1$ term is the dummy variable for treatments. Terms $x_2,x_3,x_4$ are dummy variables for blocks


## b)

```{r}
Ex1=c(40,60,70,80)
Ex2=c(55,69,75,90)

boxplot(cbind(Ex1,Ex2))

t.test(Ex1,Ex2,paired=T)
```
There is evidence of a difference in means at the 95% confidence level
## c)

```{r}
Exams=c(Ex1,Ex2)
Treatments=rep(c("T1","T2"),c(4,4))
Blocks=rep(c("1","2","3","4"),2)
Exams
Treatments
Blocks

#we want to test if treatment means are equal

model_c=lm(Exams~Treatments+Blocks)
model_r=lm(Exams~Blocks)
```


## d)

```{r}
anova(model_r,model_c)
```


The p-value from the partial F statistic tells us that we can say with 95% confidence that the treatment means are different. Meaning we can say with confidence that the mean for exam1 is different than the mean for exam2

## e)


$T_{calc}=-4.7411$

$F_{calc}=22.478$


show numerically

```{r}
-4.7411/22.478
```


$$T_{calc}=-0.21F_{calc}$$



IGNORE BELOW

$$F_{calc}=\frac{\frac{SST}{k}}{\frac{SSE}{n-(k+1)}}$$


$$SSE=\sum (Y_i-\hat{Y_i})^2$$

$$SST=\sum n_i (\bar{y_i}-\bar{y})^2$$

$$F=\frac{\frac{\chi^2}{\nu}}{\frac{\chi^2}{\nu}}$$


$$T=\frac{Z}{\sqrt{\frac{\chi^2}{\nu}}}$$
