---
title: "Exam 1 R code 2019"
author: "Collin Dabbieri"
date: "October 4, 2019"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
dird="./"
#my function to read data 
myread=function(xls){
  fl=paste0(dird,xls)
  readxl::read_xls(fl)
}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

All answers to Q 1 part e) and all answers to Q 4 a) - e) must be placed in this document -- knitted and then uploaded to CANVAS.
All files must be in the same working directory. The s20x package will be helpful for both questions. 

# Question 1 part e)

```{r}
library(readxl)
irr=read_excel("IRRIGATE.XLS")
head(irr)

m=as.table(c(17,27,22,15,19))
m
```

## a)

```{r}
chisq.test(m,correct=FALSE)
```

p-value=0.3546
## b)

There is not sufficient evidence to rule against the NULL hypothesis that all of the underlying proportions are $\frac{1}{5}$.

## c)

```{r}
mypipj = function(tab,ni,nj,alpha=0.12) #tab = table
{
  total_points=sum(tab)
  ci_matrix_lower=array(dim=c(length(tab),length(tab)))
  ci_matrix_upper=array(dim=c(length(tab),length(tab)))
  pt_est_matrix=array(dim=c(length(tab),length(tab)))
  z=qnorm(1-alpha/2,0,1)
  for (i in 1:length(tab)){
    for (j in 1:length(tab)){
      if (i==j){
        
      }
      else{
        pihat=tab[i]/total_points
        pjhat=tab[j]/total_points
        pt_est=pihat-pjhat
        pt_est_matrix[i,j]=pt_est
      
        radicand=(pihat*(1-pihat)+pjhat*(1-pjhat)+2*pihat*pjhat)/(total_points)
        lower=pt_est-z*sqrt(radicand)
        upper=pt_est+z*sqrt(radicand)
        ci=c(lower,upper)
        ci_matrix_lower[i,j]=lower
        ci_matrix_upper[i,j]=upper
        
      }

      
    }
    
  }
  
  dimnames(ci_matrix_lower) = list( c("A", "B","C","D","E"),c("A","B","C","D","E"))
  dimnames(ci_matrix_upper) = list( c("A", "B","C","D","E"),c("A","B","C","D","E")) 
  dimnames(pt_est_matrix) = list( c("A", "B","C","D","E"),c("A","B","C","D","E"))

  
  
  # Now create colored barplot with hypothesis values and sample values
  num_categories=length(tab)
  hypoth_val=1./num_categories
 
  
  true_vals=c()
  for (i in 1:num_categories){
    ptemp=tab[i]/total_points
    true_vals=append(true_vals,ptemp)
  }

  
  barplot(rbind(true_vals,hypoth_val),beside=TRUE,legend = c("sample", "hypothesis"),xlab="Categories",ylab="Proportion",col=c('lightblue','magenta'))
  
  return(list(ci_lower_bound=ci_matrix_lower,ci_upper_bound=ci_matrix_upper,pt_est=pt_est_matrix))
  
}

mypipj(m,ni=27,nj=15)
```



## d)

Confidence interval for $p_2-p_4=(0.0209,0.2190)$
# Question 4

```{r}
bid=myread("BIDMAINT.XLS")
#head(bid)
bid = within(bid,{
  y<-COST
  x1<-as.integer(X1)
  x2<-as.integer(X2)
})
bid
```



## a)

```{r}
model=lm(bid$y~bid$X1+bid$X2)
summary(model)
```
Yes, the p-value for the F test is less than 0.05, therefore we can reject the NULL.

## b)

```{r}
library(s20x)
ci=ciReg(model)
```


## c)
Yes, p-value for F test is <0.05, model is adequate

## d)
168.9
## e)
0.04515
